lifespan_eda

### 2024-02-17 eda/lifespan_eda.Rmd
rm subjects with euler >2 MADs above study median

Load data from Jakob
```{r}
cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly.csv"), stringsAsFactors = TRUE)

@@

nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
```

@@

@@

@@

```{r}
cn.df.site <- cn.df.unique %>% 
  dplyr::filter(site != "NA") %>%
  mutate(study = droplevels(study),
         site = gsub(" ", "_", trimws(site))) %>%
  unite(site, study, site, sep = "_", remove = FALSE) %>% #so sites with the same name don't get combined across studies!
  na.omit()
n_unique(cn.df.site$participant)
n_unique(cn.df.site$participant) == nrow(cn.df.site)

@@

```{r}
fwrite(cn.df.site, file=paste0(base_path, "/data/lifespan_CN_site-level.csv"))
```

@@

```{r}
#clean up site NAs
cn.df.imp <- cn.df.unique %>%
  mutate(site = trimws(as.character(site)),
         study = trimws(study)) %>%
  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="_")))) %>%
  na.omit()

@@

```{r}
fwrite(cn.df.imp, file=paste0(base_path, "/data/lifespan_CN_imp-sites.csv"))
```

>>>>

Load data from Jakob - updated with euler numbers!
```{r}
cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)

@@

nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
```
Write out data for andrew
```{r, eval=FALSE}
andrew.df <- cn.df.unique %>%
  dplyr::filter(study != "UKB" & age_days >= (365.25*18))
nrow(andrew.df)
fwrite(andrew.df, file=paste0(base_path, "/data/lifespan_CN_adult.csv"))
```

@@

euler # noted as surfholes

@@

## Curate

Visualizing euler distributions
```{r}
cn.df.unique %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=surfholes))
```

Removing subjects w/o surfholes, subjects with surfholes values > 2 MADs above study median (same as sensitivity analyses in LBCC)

```{r}
cn.df.eulerfilt <- cn.df.unique %>%
  dplyr::filter(!is.na(surfholes))

study_mad <- cn.df.eulerfilt %>%
  group_by(study) %>%
  summarize(study_euler_mad = mad(surfholes),
            # study_euler_madconst1 = mad(surfholes, constant=1),
            study_euler_median = median(surfholes)) %>%
  mutate(euler_threshold = study_euler_median+(2*study_euler_mad))

left_join(cn.df.eulerfilt, study_mad) %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=surfholes)) +
  geom_point(aes(x=study, y=(study_euler_median+2*study_euler_mad)), color="red")

left_join(cn.df.eulerfilt, study_mad) %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=surfholes)) +
  geom_point(aes(x=study, y=(study_euler_median+2*study_euler_madconst1)), color="red")
```

using default mad() constant, looks better

```{r}
cn.df.eulerfilt2 <- left_join(cn.df.eulerfilt, study_mad) %>%
  dplyr::filter(surfholes <= euler_threshold)
```

Removed `r nrow(cn.df.unique) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain

@@

```{r}
cn.df.site <- cn.df.eulerfilt2 %>% 
  dplyr::filter(site != "NA") %>%
  mutate(study = droplevels(study),
         site = gsub(" ", "_", trimws(site))) %>%
  unite(site, study, site, sep = "_", remove = FALSE) %>% #so sites with the same name don't get combined across studies!
  na.omit()
n_unique(cn.df.site$participant)
n_unique(cn.df.site$participant) == nrow(cn.df.site)

@@

```{r}
fwrite(cn.df.site, file=paste0(base_path, "/data/lifespan_CN_site-level_euler.csv"))
```
write out imagen data for testing
```{r}
imagen.df <- cn.df.site %>%
  dplyr::filter(study=="IMAGEN")
# skim(imagen.df)

fwrite(imagen.df, file=paste0(base_path, "/data/imagen_CN.csv"))
```

troubleshooting combat

```{r, eval=FALSE}
gamlss(formula = rh_Vol_posteriorcingulate ~ pb(age_days) + sexMale + pb(sex.age) + site, sigma.formula= ~ pb(age_days) + sexMale + site, data=imagen.df)
```

@@

```{r}
#clean up site NAs
cn.df.imp <- cn.df.eulerfilt2 %>%
  mutate(site = trimws(as.character(site)),
         study = trimws(study)) %>%
  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="_")))) %>%
  na.omit()

@@

```{r}
fwrite(cn.df.imp, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler.csv"))
```
